---
output: 
  github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# RHL30

An R Package for the RHL30 prognostic predictor in relapsed Hodgkin lymphoma
samples. The predictor was published at:

Chan, F. C. et al. Prognostic Model to Predict Post-Autologous Stem-Cell 
Transplantation Outcomes in Classical Hodgkin Lymphoma. J Clin Oncol 
JCO2017727925 (2017) doi:10.1200/JCO.2017.72.7925.

# How to Install

To get the released version from CRAN:

```{r, eval = FALSE}
install.packages("survutils")
```

You can also get survutils through conda:

```{bash, eval = FALSE}
conda install -c fongchun r-survutils
```

To install the latest developmental version from github:

```{r, eval = FALSE}
devtools::install_github("tinyheero/survutils")
```

# How to use

We will be use the BCCA relapsed hodgkin lymphoma cohort (n=68) from the paper
as an example of how to generate RHL30 predictor score. The following steps will
reproduce the scores from the paper. 

First, let's load the RHL30 package and the RHL30 model:

```{r}
library("RHL30")
rhl30_model_df <- get_rhl30_model_coef_df()
rhl30_model_df
```


```{r chung_tag}
hk_genes <- 
  dplyr::filter(rhl30_model_df, gene_type == "housekeeper") %>%
  dplyr::pull("gene_name")

exprs_file <- 
  system.file("extdata", "bcca_rhl_rhl30_gene_exprs_mat.tsv", package = "RHL30")
exprs_mat <- load_exprs_mat(exprs_file)
sample_normalizer_values <- get_sample_normalizer_value(exprs_mat, hk_genes)
exprs_mat_norm <- normalize_exprs_mat(exprs_mat, sample_normalizer_values)
rhl30_df <- get_rhl30_scores_df(exprs_mat_norm, rhl30_model_df)
```

# R Session

```{r}
devtools::session_info()
```
