---
output: 
  github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r echo = FALSE}
library("knitr") 

opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# RHL30

An R Package for the RHL30 prognostic predictor in relapsed Hodgkin lymphoma
(RHL) samples. The predictor was published at:

Chan, F. C. et al. Prognostic Model to Predict Post-Autologous Stem-Cell 
Transplantation Outcomes in Classical Hodgkin Lymphoma. J Clin Oncol 
JCO2017727925 (2017) doi:10.1200/JCO.2017.72.7925.

# How to Install

To install this package, you need to first have the package `devtools` 
installed, then you run:

```{r, eval = FALSE}
devtools::install_github("tinyheero/RHL30")
```

# How to use

We will be use the BCCA RHL cohort (n=68) from the paper as an example of how to 
generate RHL30 predictor score. The following steps will reproduce the RHL30 
scores from the paper. 

First, let's load the RHL30 package and the RHL30 model:

```{r}
library("RHL30")
library("dplyr")
rhl30_model_df <- get_rhl30_model_coef_df()
rhl30_model_df
```

The model contains a total of 30 genes:

* 18 genes that make the model
* 12 housekeeper genes that are used to normalize the data

Let's load the expression data of the BCCA RHL cohort that is provided as part
of this package.

```{r}
hk_genes <- 
  filter(rhl30_model_df, gene_type == "housekeeper") %>%
  pull("gene_name")

exprs_file <- 
  system.file("extdata", "bcca_rhl_rhl30_gene_exprs_mat.tsv", package = "RHL30")
exprs_mat <- load_exprs_mat(exprs_file)
dim(exprs_mat)
```

The expression data contains the 30 genes (rows) and 68 samples (columns). Next 
we calculate the normalizer values for each sample:

```{r}
sample_normalizer_values <- get_sample_normalizer_value(exprs_mat, hk_genes)
```

In the paper, a threshold of 35 was set to exclude poor quality samples. This 
was done very low normalizer values often lead to very high normalized 
expression values. We can apply this threshold to eliminate poor quality 
samples:

```{r}
high_quality_samples <- 
  names(sample_normalizer_values[sample_normalizer_values > 35])
filtered_exprs_mat <- exprs_mat[, high_quality_samples]
dim(filtered_exprs_mat)
```

This eliminates 2 poor quality samples leaving us with 66 samples. **Note that
the sample HL1120 did not receive ASCT and thus was not reported in figure 4 of
the paper. As such, the final number in figure 4 is 65 samples.**

Let's normalize our expression matrix and generate the RHL30 scores for each
sample:

```{r}
filtered_exprs_mat_norm <- 
  normalize_exprs_mat(filtered_exprs_mat, sample_normalizer_values)
rhl30_df <- get_rhl30_scores_df(filtered_exprs_mat_norm, rhl30_model_df)
head(rhl30_df)
```

# R Session

```{r}
devtools::session_info()
```
